---
title: "To be determined"
author:
  - name: Paulo Martins r2015469
    affil: 1
  - name: Beatriz Fonseca r20201599
    affil: 1
affiliation:
  - num: 1
    address: Nova Information Management School
column_numbers: 4
output: 
  posterdown::posterdown_html:
    self_contained: false
poster_width: "55.1in" # Double check size against standard specifications.
poster_height: "39in"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```
# **Introduction**
```{r, include=FALSE}
df <- readxl::read_excel("C:") # Consider dynamic pathing
```
```{r echo=FALSE, out.width='100%', fig.align='center', fig.cap='Figure 1. Template 1'}
```
# **Goals**
This research project sets out to study the relationship between import tariffs and vehicle import, to and from the U.S., Chinese, and E.U. Markets, and therefore measure the effectivness of said public policy.

```{r echo=FALSE, out.width='100%', fig.align='center', fig.cap='Figure 2. Figure 2 template.'}
```
# **Methodology**
TODO: Determined the methodologies to be used.
```{r echo=FALSE, out.width='100%', fig.align='center', fig.cap='Figure 3. Figure 3 template.'}
```
# **Results**
The results obtained should always be presented as a confidence interval.
```{r echo=FALSE}
```
```{r echo=FALSE}
```
```{r echo=FALSE, fig.cap='Table 1. '}
```
```{r echo=FALSE, out.width='100%', fig.align='center', fig.cap='Figure 4. Figure 4 template'}
```
# **Conclusions**
```{r echo=FALSE, out.width='100%', fig.align='center', fig.cap='Figure 5. Figure 5 template'}
```
# **References**
Hyndman, R., & Athanasopoulos, G. (2021). Forecasting: Principles and Practice. (3rd ed.) OTexts. <https://otexts.com/fpp3/>
SHAPIRO, S. S., & WILK, M. B. (1965). An analysis of variance test for normality (complete samples). Biometrika, 52(3--4), 591--611. <https://doi.org/10.1093/biomet/52.3-4.591>

# **Footnotes**

